---
title: "**R_WORKSHOP**"
author: "**DIR.T@ RWILLS_STATS**"
date: "2/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#   **HARCLUX ENTERPRISE**


##           *WELCOME WILFRED TO DATA GURUS*



 #             **Packages & DATA**
 
### Packages
```{r packages, message=FALSE}
#set your working directory
setwd("~/RSTUDIO FILES/R-WORKSHOP")

#install.packages("titdyverse")
library(tidyverse)

# install.packages("devtools")
#devtools::install_github("Shelmith-Kariuki/rChambua")

##package still under development not yet in CRAN
library(rChambua)
```

### Data

Reading in your data
```{r data}


#from the rchambua pckg
##df <- rChambua::wafanyikazi

#raw file
df <- read.csv("Wafanyikazi.csv",
               stringsAsFactors = F,
               header = T)

#checking for column names
colnames(df)

#checking for dimension
dim(df)

#checking the strucure of our data
str(df)
```




##            **TIDYVERSE** 
The following are must know functions in Tidyverse package as they will form your day to day data wrangling/management codes.


### SELECT 

using the select ()to pick variables


```{r select}
#selecting Age and gender variables

age_gender <- df %>%
  select(Age, Gender)

head(age_gender)

#checking for variables starting with letter "a"
starts_a <- df %>%
  select(starts_with("A"))

head(starts_a)

#contains letter "a"
contains_a <- df %>%
  select(contains("A"))

head(contains_a)

#end with ends with letter "a"
ends_a <- df %>%
  select(ends_with("t"))


head(ends_a)
```


### FILTERING

using the filter() helps us to choose cases where conditions are true.

```{r filter}

#filter females only
female_data <- df %>%
  filter(Gender == "Female")

head(female_data)



#filter age 30 and below
age_30 <- df %>%
  filter(Age <= 30)

head(age_30)


#filter promotion
female_data <- df %>%
  filter(Gender == "Female")
head(female_data)


#filter females nyeri age< 26 from nyeri county earn <5000
female_nyeri <- df %>%
  filter(Gender == "Female" & County == "Nyeri" &
           Age >= 26 & Income < 5000)

head(female_nyeri)


#filter females and exclude the juniors

females_NoJnr <- df %>%
  filter(Gender == "Female" & Income > 6000 & Role != "Junior")

head(females_NoJnr)
```
### MUTATE 

using the mutate() helps us to create variables


```{r mutate}


income_df <- df %>%
  select(Income) %>%
  mutate(Income2 = Income * 100)

head(income_df)



#creating a swahili gender col
swahili <- df %>%
  select(Gender) %>%
  mutate(jinsia =
           if_else(Gender == "Male", "Mme", "Mke"))

head(swahili)
```



#### IFELSE 

using the ifelse() helps us to shift from one condition to another

```{r ifelse}


#Nested Ifelse
##Department
dep_ifesle <- df %>%
  select(Department) %>%
  mutate(Department2 =
           ifelse(
             Department == "Finance",
             'F',
             ifelse(
               Department == "Research Analyst" ,
               "R",
               ifelse(
                 Department == "Operations",
                 "O",
                 ifelse(Department == "Associate", "A", "D")
               )
             )
           ))

head(dep_ifesle)



#creating age Groups
age_group <- df %>%
  select(Age) %>%
  mutate(age2 =
           ifelse(Age < 25, "Young",
                  ifelse(Age > 25 & Age < 40, "Middle", "old")))

head(age_group)
```




### Renaming 
using the rename() helps us to rename variables


```{r rename}

#rename age to Miaka(swahili)
age_miaka <- df %>%
  rename(miaka = Age)

head(age_miaka)
```


### Group_by & Summarise

Group_by works close with summarise they help us cluster common variable categories and get say statistical summaries of the same.

```{r group_summ}

##Group according to Gender and get the summary stats

gender_group <- df %>%
  group_by(Gender) %>%
  summarise(mean_age = mean(Age)) 

head(gender_group)




#group according to Department and get the summary stats

Depart_group <- df %>%
  group_by(Department) %>%
  summarise(mean_age = mean(Age))

head(Depart_group)



#group according to county and get the summary stats

county_group <- df %>%
  group_by(County) %>%
  summarise(
    minimum = min(Income),
    max = max(Income),
    mean = mean(Income)
  )

head(county_group)
```

# **WELCOME WILFRED TO DATA GURUS TEAM**
```{r WILFRED}


Wilfred <- df %>% 
  filter(Gender=="Male" &  Income>=5000) 
  
Wilfred %>% 
  group_by(Role,County, Age, Department) %>% 
  summarise(salary=
              mean(Income)) %>% 
  ggplot(aes(Age, salary))+
  geom_point(aes(color=Role)) +
  facet_wrap(~Department)
```

